\begin{abstract}
  Mutation testing is widely used in research (even if not in
  practice).  Mutation testing tools usually target only one
  programming language, and rely on parsing a program to generate
  mutants, or operate not at the source level but on compiled
  bytecode.  Unfortunately, developing a robust mutation testing tool
  for a new language in this paradigm is a difficult and
  time-consuming undertaking.  Moreover, bytecode/intermediate
  language mutants are difficult for programmers to read and
  understand.  This paper presents a simple tool, called {\tt
    universalmutator}, based on regular-expression-defined transformations of
  source code.  The primary drawback of such an approach is that our
  tool can generate invalid mutants that do not compile, and sometimes
  fails to generate mutants that a parser-based tool would have
  produced. Additionally, it is incompatible with some approaches to
  improving the efficiency of mutation testing.  However, the
  regexp-based approach provides multiple compensating advantages.
  First, our tool is easy to adapt to new languages; e.g., we present
  here the first mutation tool for Apple's Swift programming
  language.  Second, the method makes handling multi-language programs
  and systems simple, because the same tool can support every
  language.  Finally, our approach makes it easy for users to add
  custom, project-specific mutations.
{\bf VIDEO AT: \url{https://youtu.be/2T23mPVq2Hs}}
\end{abstract}

% LocalWords:  bytecode universalmutator multi
